Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India

### 2019-06-27 draft.Rmd
Uppangala 0.1.9054 Francois sharing

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt, Valérie Raevel, Maxime Réjou-Méchain, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2: default
linestretch: 1.5
csl: bibliography/mee.csl
bibliography: ["bibliography/refs.bib", "bibliography/library.bib"] 
link-citations: yes
---

@@

```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values. Marginal plots represent the density distribution of species on the two principal component axis for each strata. See table \\@ref(tab:Traits) for abbreviation.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
spectrum.density <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(WES = Comp.1, LES = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(value, color = spectrum, fill = spectrum)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~strata, nrow = 2) +
  coord_flip() +
  scale_fill_manual("PCA axis",
                    labels = c("WES" = "PCA1", "LES" = "PCA2"),
                    values = c("green", "purple")) +
  scale_color_manual("PCA axis",
                     labels = c("WES" = "PCA1", "LES" = "PCA2"),
                     values = c("green", "purple"))
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2)
# cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")), 
       file = "PCA.png",
       path = './data/',
       width = 200, height = 150, unit = 'mm',
       dpi = 300)
```

```{r varianceTest}
data <- pca.plot$data %>% 
  select(Comp.1, Comp.2, strata) %>% 
  # rename(WES = Comp.1, LES = Comp.2) %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
  melt(id.vars = "strata", variable.name = "spectrum")
Variance <- list(
  LES = list(
    C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
    U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
 WES = list(
   C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
   U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
)
Leven_test <- list(
  C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
  U = round(car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1], 3)
)
```

@@

```{r GWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  # select(-community, -LES, -WES) %>% 
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  # facet_wrap(~trait, scales = "free", nrow = 1) +
  facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
```

@@

```{r LM}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  mutate(community = as.character(community)) %>% 
  left_join(R, by = "community") %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                     direction = 'both', trace = 0)) %>% 
  tidy(reg) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
  mutate_all(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      mutate(communtiy = as.character(community)) %>% 
      left_join(R, by = "community") %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared), by = c("strata", "trait")
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), and canopy height (Canopy), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
```

@@

```{r CEeffect}
# L %>% 
#   melt(id.vars = "com",
#        variable.name = "SpCode",
#        value.name = "individuals") %>% 
#   filter(individuals > 0) %>% 
#   left_join(Q %>% 
#               select(SpCode, strata)) %>% 
#   filter(!is.na(strata)) %>% 
#   group_by(com, strata) %>% 
#   summarise(species = length(unique(SpCode)),
#             individuals = sum(individuals)) %>% 
#   ungroup() %>% 
#   # group_by(strata) %>% 
#   summarise(species = mean(species), individuals = mean(individuals))
# genPFT() %>%
#   mutate(CE = gsub('([[:alpha:]])', "", CE)) %>%
#   filter(!is.na(CE)) %>%
#   mutate(LA = log(LA)) %>%
#   mutate(SLA = log(SLA)) %>%
#   select(CE, Sp, WD, LT, LA, LDMC, LTD, SLA) %>%
#   reshape2::melt(id.vars = c("CE", "Sp"), variable.name = "Trait") %>%
#   group_by(Trait) %>%
#   do(anova = TukeyHSD(aov(lm(value ~ CE, data = .)))) %>%
#   tidy(anova) %>%
#   filter(adj.p.value < 0.05)
my_comparisons <- list( c("1", "2"), c("2", "3"), c("3", "4"),
                        c("4", "5"), c("1", "3"), c("3", "5"),
                        c("1", "5"),  c("2", "4"), c("2", "5"))
genPFT() %>% 
  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
  filter(!is.na(CE)) %>% 
  select(CE, WD, LT, LA, LDMC, LTD, SLA) %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  reshape2::melt(id.vars = "CE", variable.name = "Trait") %>% 
  ggplot(aes(CE, value, fill = CE)) +
  geom_boxplot() +
  facet_wrap(~Trait, scales = "free") +
  ggpubr::stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..)) +
  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
  ggsave(file = "CE.png",
         path = './data/',
         width = 200, height = 200, unit = 'mm',
         dpi = 300)

genPFT() %>% 
  select(CE) %>% 
  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
  group_by(CE) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(P = round(n/sum(n)*100))
```

>>>>

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt, Valérie Raevel, Maxime Réjou-Méchain, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
date: '`r Sys.Date()`'
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
linestretch: 1.5
csl: bibliography/mee.csl
bibliography: ["bibliography/refs.bib", "bibliography/library.bib"] 
link-citations: yes
---

@@

```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values. Marginal plots represent the density distribution of species on the two principal component axis for each strata. See table \\@ref(tab:Traits) for abbreviation.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
       file = "PCA.png",
       path = './data/',
       width = 200, height = 150, unit = 'mm',
       dpi = 300)
```

```{r varianceTest}
data <- pca.plot$data %>% 
  select(Comp.1, Comp.2, strata) %>% 
  # rename(WES = Comp.1, LES = Comp.2) %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
  melt(id.vars = "strata", variable.name = "spectrum")
Variance <- list(
  LES = list(
    C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
    U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
 WES = list(
   C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
   U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
)
Leven_test <- list(
  C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
  U = round(car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1], 3)
)
pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>% 
  group_by(spectrum) %>% 
  do(anova = aov(value ~ strata, data = .)) %>% 
  broom::tidy(anova)
```

@@

```{r GWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>%
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  # select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  # facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3)))) +
  ggsave(file = "boxplot.png",
         path = './data/',
         width = 200, height = 150, unit = 'mm',
         dpi = 300)
```

@@

```{r LM}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  mutate(community = as.character(community)) %>% 
  left_join(R, by = "community") %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
     direction = 'both', trace = 0)) %>% 
  # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
  #    direction = 'both', trace = 0) %>%
  tidy(reg) %>%
  group_by(strata, trait, term) %>%
  # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          # stars(p.value, ns = '   '))) %>% 
  transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
  mutate_all(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      mutate(communtiy = as.character(community)) %>% 
      left_join(R, by = "community") %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared), by = c("strata", "trait")
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  # bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), and canopy height (Canopy), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations.', format = 'pandoc')
```

@@

```{r CEeffect}
# L %>% 
#   melt(id.vars = "com",
#        variable.name = "SpCode",
#        value.name = "individuals") %>% 
#   filter(individuals > 0) %>% 
#   left_join(Q %>% 
#               select(SpCode, strata)) %>% 
#   filter(!is.na(strata)) %>% 
#   group_by(com, strata) %>% 
#   summarise(species = length(unique(SpCode)),
#             individuals = sum(individuals)) %>% 
#   ungroup() %>% 
#   # group_by(strata) %>% 
#   summarise(species = mean(species), individuals = mean(individuals))
# genPFT() %>%
#   mutate(CE = gsub('([[:alpha:]])', "", CE)) %>%
#   filter(!is.na(CE)) %>%
#   mutate(LA = log(LA)) %>%
#   mutate(SLA = log(SLA)) %>%
#   select(CE, Sp, WD, LT, LA, LDMC, LTD, SLA) %>%
#   reshape2::melt(id.vars = c("CE", "Sp"), variable.name = "Trait") %>%
#   group_by(Trait) %>%
#   do(anova = TukeyHSD(aov(lm(value ~ CE, data = .)))) %>%
#   tidy(anova) %>%
#   filter(adj.p.value < 0.05)
# genPFT() %>%
#   mutate(CE = gsub('([[:alpha:]])', "", CE)) %>%
#   filter(!is.na(CE)) %>%
#   mutate(LA = log(LA)) %>%
#   mutate(SLA = log(SLA)) %>%
#   select(CE, Sp, WD, LT, LA, LDMC, LTD, SLA) %>%
#   reshape2::melt(id.vars = c("CE", "Sp"), variable.name = "Trait") %>%
#   mutate(CE = as.numeric(CE)) %>% 
#   group_by(Trait) %>%
#   do(lm = lm(value ~ CE + Sp, data = .)) %>%
#   tidy(lm) %>% 
#   filter(term == "CE") %>% 
#   filter(p.value < 0.05)
my_comparisons <- list( c("1", "2"), c("2", "3"), c("3", "4"),
                        c("4", "5"), c("1", "3"), c("3", "5"),
                        c("1", "5"),  c("2", "4"), c("2", "5"))
genPFT() %>% 
  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
  filter(!is.na(CE)) %>% 
  select(CE, WD, LT, LA, LDMC, LTD, SLA) %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  reshape2::melt(id.vars = "CE", variable.name = "Trait") %>% 
  ggplot(aes(CE, value, fill = CE)) +
  geom_boxplot() +
  facet_wrap(~Trait, scales = "free") +
  ggpubr::stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..)) +
  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
  ggsave(file = "CE.png",
         path = './data/',
         width = 200, height = 200, unit = 'mm',
         dpi = 300)
genPFT() %>% 
  select(CE) %>% 
  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
  group_by(CE) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(P = round(n/sum(n)*100))
genPFT() %>%
  mutate(CE = as.numeric(gsub('([[:alpha:]])', "", CE))) %>%
  filter(CE != 5) %>% 
  filter(!is.na(CE)) %>%
  ggplot(aes(CE, WD, group = NA)) +
  geom_boxplot(aes(group = CE), col = "lightgrey") +
  geom_point(aes(col = Sp), alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = "none")
```
